---
title: "Dottable Homework"
author: "Collin Brown and Jesse Brandt"
date: "2025-03-04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(dplyr)
library(data.table)
```

# Load data
```{r}
nycdata <- fread("nycdata.csv")
glimpse(nycdata)
```

# Problem 1
Use and show data.table code to select the variables year, month, day, and hour from the imported flights data
```{r}
nycdata[, .(year,month,day,hour)]
```

# Problem 2
Use and show data. table code to produce a table that shows a carrier of DL, an origin of JFK and a destination of SEA
```{r}
nycdata[carrier == "DL" & origin == "JFK" & dest == "SEA"]
```

# Problem 3
Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
nycdata[carrier == "UA" & month == 3 & air_time < 330]
```

# Problem 4
Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
nycdata_4 <- nycdata |>
  filter(carrier == "UA", month == 3, air_time < 330)
nycdata_4
```

# Problem 5
Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour.
```{r}
nycdata_5 <- nycdata
nycdata_5[, speed := (distance / air_time * 60)]
nycdata_5
```

# Problem 6
Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.
```{r}
nycdata_6 <- nycdata |>
  mutate(speed = (distance / air_time) * 60)
nycdata_6
```

# Problem 7a
Show and use coding to change the carrier abbreviation of UA to UniitedAir (sic),
7a. data.table method
```{r}
nycdata_7a <- nycdata
nycdata_7a[carrier == "UA", carrier := "UnitedAir"] 
unique(nycdata_7a$carrier) # list all carriers
nycdata_7a
```

# Problem 7b
Show and use coding to change the carrier abbreviation of UA to UniitedAir (sic),
7b. tidyverse method (Use a sequence of dplyr commands so that you can see the change in your table)
```{r}
nycdata_7b <- nycdata |>
  mutate(carrier = case_when(carrier == "UA" ~ "UnitedAir", TRUE ~ carrier))
unique(nycdata_7b$carrier) # list all carriers
nycdata_7b
```
